k-Space Domain Parallel Transmit Pulse Design

### 2020-08-17 Manuscript/Theory.tex
wag edited first section of theory

\subsection*{k-space domain parallel-transmit pulse design}
\par Here we describe the derivation of the highly parallelizable k-sapce domain design method. We propose to find a system matrix $\mathbf{W}$ that directly maps a desired excitation pattern in k-space $\mathcal{F}(\mathbf{d})$, which is the Fourier transform of the spatial domain desired pattern $\mathbf{d}$, to the corresponding desired RF pulse $\mathbf{b}$, as:
\begin{equation*}
\mathbf{b}=\mathbf{W}\mathcal{F}(\mathbf{d})
\end{equation*}
$\mathcal{F}(\mathbf{d})$ is the vectorized k-space excitation pattern with length of $N_s$, which is the number of spatial/k-space locations.  $\mathbf{b}$ is the concatenated multi-channel RF samples with length of $N_tN_c$, where $N_t$ and $N_c$ are the numbers of time points and coils, respectively. $\mathbf{W}$ is the $N_tN_c$ by $N_s$ system matrix, which will latter be shown sparse.
Therefore, once the $\mathbf{W}$ matrix is found, the RF design problem can be instantaneously solved by a sparse matrix multiplication, and there is no matrix inversion or iterative CG method needed. Furthermore, in design problems involved iterative procedures, such as the phase updates in magnitude-least-squares pulse designs \cite{setsompop2008magnitude} and the regularization factor updates in power/roughness constrained pulse design, the sparsity of the $\mathbf{W}$ matrix makes the iterative process much more efficient in terms of memory and speed. In the work of Katscher et al \cite{katscher2003transmit}, parallel transmit pulse design was also formulated in the k-space domain. However, in Ref \cite{katscher2003transmit}, the matrix $\mathbf{W}$ was solved as a whole, which required constructing and inverting a large $\mathbf{S}^H\mathbf{S}$ matrix, which will be explained in the following text. Our proposed method finely divides the process of finding $\mathbf{W}$ into many independent instances, which allows the use of parallel computing to largely accelerate the computation. Furthermore, the $\mathbf{S}^H\mathbf{S}$ matrices of different instances are each small, and our proposed method efficiently constructs these matrices, which further accelerates the computation.

\par We can solve for the columns of the $\mathbf{W}$ matrix independently to each other, based on the key observation that if the desired k-space pattern is a delta function at a given k-space location $\vec{k}_i$ ($1\leq i \leq N_s$), then only the $i$-th column $\mathbf{w}_i$ of $\mathbf{W}$ relates the desired k-space pattern to the desired RF pulse. In this case, the vector $\mathbf{w}_i$ is identical to the desired RF. Therefore, the question of finding the $i$-th column $\mathbf{w}_i$ of $\mathbf{W}$ can be restated as the following: what should the RF pulse samples be in order to generate a unit delta function at the location $\vec{k}_i$ in k-space, and zeros elsewhere. Mathematically, the relationship that should be satisfied by the weights in column $\mathbf{w}_i$  can be expressed as \cite{katscher2003transmit}:   
\begin{equation}\label{eq:ScalerForm}
\mathbf{\delta}(\vec{k}-\vec{k}_i)=\sum_{t=1}^{N_t}\sum_{c=1}^{N_c}w_c(\vec{k}_t)\mathbf{s}_c(\vec{k}-\vec{k}_t)
\end{equation}
where $t$ indexes RF samples on the pulse’s excitation k-space trajectory, and $c$ indexes coils. $w_c(\vec{k}_t)$ is the $((c-1)Nt+t)$-th element of the $\mathbf{w}_i$ vector. $s_c(\vec{k}-\vec{k}_t)$ is the Fourier transform of coil-$c$'s $B_1^+$ map , shifted to be centered at excitation k-space location $\vec{k}_t$. 
This relationship can be restated in a matrix-vector form as:                                       
\begin{equation}\label{eq:MatrixForm}
\mathbf{Sw}_i=\mathbf{\delta}_i
\end{equation}
and the regularized pseudoinverse solution for the weight $\mathbf{w}_i$ is                       
\begin{equation}\label{eq:Solution}
\mathbf{w}_i=\left( \mathbf{S}^{H} \mathbf{S} + \lambda \mathbf{I} \right) ^{-1} \mathbf{s}_i^{H}
\end{equation}
where $\mathbf{S}$ is the $N_s$ by $N_cN_t$ sensitivity matrix consisting of horizontally stacked $\mathbf{s}_c(\vec{k}-\vec{k}_t)$ vectors for all $c$ and $t$, and $\mathbf{s}_i^{H}$ is the conjugate transpose of the $i$-th row of $\mathbf{S}$. Since k-space sensitivity maps are localized around DC, only few elements in $\mathbf{s}_i^{H}$ and $\mathbf{w}_i$ corresponding to $\vec{k}_t$s near $\vec{k}_i$ are non-zero. In order words, only few RF samples on the excitation trajectory near the target $\vec{k}_i$ can and should contribute energy to the unit delta function at this location, and all other RF samples on the excitation trajectory should automatically be zero. Therefore, in Equation \ref{eq:ScalerForm}, only few subscript $t$s are needed in the summation, and in Equation \ref{eq:MatrixForm} and \ref{eq:Solution}, the size of matrix $\mathbf{S}$ is much smaller than it was defined in Ref \cite{katscher2003transmit}. It also indicates the system matrix $\mathbf{W}$ is sparse. 
Techniques for efficient parallelization of the computation of the column $\mathbf{w}_i$s and efficient construction of the $\mathbf{S}^{H}\mathbf{S}$ matrices are also developed and will be discussed in the following text.

@@

\par A key advantage of the proposed pulse design formalism is that it can be broken down into many small problems and finely parallelized, because the process of solving each column of $\mathbf{W}$ is independent of other columns. The columns can be solved fully independently, or patchwise for target locations in neighborhoods, wherein all points in the neighborhood share the same $\mathbf{S}^{H}\mathbf{S}$ matrix. 

@@

An example of this concept is shown in Figure \ref{fig:Patch}. For simplicity of visualization, it is demonstrated in a 2D problem, but the concept is easily extended to 3D. The figure depicts the solution of the $\mathbf{W}$ matrix 16 columns at a time, whose corresponding delta functions live within a patch of width 4 cycle/FOV (yellow points in Figure \ref{fig:Patch}). In this instance where the inclusion width is also 4 cycle/FOV, we truncate the $B_1^+$ map Fourier transforms to zero outside a 4 by 4 cycle/FOV square centered at DC. As a result, we must consider all the excitation trajectory points within a 12 by 12 cycle/FOV square, which is 4 cycle/FOV extended outside of the patch width 4 square, since they may contribute energy to the target locations corresponding to the 16 $\mathbf{W}$ matrix columns. For a single instance of the 16-point target patch shown in Figure \ref{fig:Patch}, the 46 green points on the excitation trajectory may contribute significant energy and are considered in the weight design. Furthermore, we must constrain the total excitation to zero at all k-space locations within a 20 by 20 cycle/FOV red patch in Figure \ref{fig:Patch}. This is because any energy deposited at each considered excitation trajectory points will affect not only the target points in the yellow square, but also the whole 2$\times$4 wide patch around that trajectory point. And the 20 by 20 cycle/FOV red patch is the total of the points that can be affected by all the excitation trajectory points considered in this instance. Therefore, in the solution for this instance, $\mathbf{w}$ is a 46$N_c\times$16 matrix, $\mathbf{S}$ is a 400$N_c\times$46$N_c$ matrix, $\mathbf{S}^{H}\mathbf{S}$ is a 46$N_c\times$46$N_c$ matrix, and $\mathbf{s}_i^{H}$ is a 46$N_c\times$16 matrix. 
%Since $\mathbf{s}_i^{H}$ is no longer related to the $i$th row of $\mathbf{S}$ and $i$th column of $\mathbf{W}$ only, it will be denoted as $\mathbf{S_{targ}}$ in later discussions. 

@@

\par The patch width determines the total number of instances to be solved. The inclusion width determines the accuracy of $B_1^+$ information utilized in the design. Together, the patch width and the inclusion width determine how many excitation trajectory points are included in each instance, which further determines the size of $\mathbf{S}^{H}\mathbf{S}$ whose construction is most computationally burdensome within an instance. Therefore, increasing patch width reduces the number of $\mathbf{S}^{H}\mathbf{S}$ matrices to be constructed but increases the burden of constructing each of them. An optimal patch width should be decided case by case for different excitation trajectories.

\subsection*{Efficient construction of $\mathbf{S}^{H}\mathbf{S}$ matrix}
\par Constructing $\mathbf{S}^{H}\mathbf{S}$ by finding $\mathbf{S}$ is computationally expensive. As described previously, the $((c-1)N_c+t)$-th column of the  $\mathbf{S}$ matrix, $\mathbf{S}_c(\vec{k}-\vec{k}_t)$, is the k-space sensitivity map of the $c$-th coil $\mathbf{S}_c(\vec{k})$ shifted to excitation k-space trajectory location $\vec{k}_t$. These shifts for each column of $\mathbf{S}$ are achieved by applying a linear phase modulation to the spatial domain $B_1^+$ map, followed by a fast Fourier transform (FFT). This process becomes computationally overwhelming when there are many RF samples, which is often the case. Furthermore, the matrix multiplication between $\mathbf{S}^{H}$ and $\mathbf{S}$ is also computationally intensive due to the large row dimension of $\mathbf{S}$ (400$Nt$ in the above example in Figure \ref{fig:Patch}).

\par Inspired by a rapid GRAPPA calibration method for arbitrary 2D/3D non-Cartesian trajectories proposed in Ref \cite{luo2019grappa}, we have developed a fast and memory-efficient algorithm to solve this problem. A key innovation of Ref \cite{luo2019grappa} is that the algorithm constructs the Hermitian $\mathbf{S}^{H}\mathbf{S}$ matrix directly by finding each of the matrix's lower triangle elements through interpolation, instead of building the $\mathbf{S}$ matrix explicitly followed by matrix multiplication. Conceptually, each element in the Hermitian matrix $\mathbf{S}^{H}\mathbf{S}$ is the vector sum of two shifted k-space sensitivity maps, which is also one single point of their convolution. With the multiplication property of Fourier transform, this convolution in k-space can be calculated from the Fourier transform of the dot product of the two spatial domain $B_1^+$ maps. For a more detailed derivation, please refer to Ref \cite{luo2019grappa}.
In practice, $\mathbf{S}^{H}\mathbf{S}$ is constructed in the following way. We first represent the $\mathbf{S}^{H}\mathbf{S}$ matrix with block matrices. Each block matrix $\mathbf{S}_i^{H}\mathbf{S}_j$ is of size $N_n$ by $N_n$, and $\mathbf{S}_i$ and $\mathbf{S}_j$ are the columns of $\mathbf{S}$ related to the k-space sensitivity maps of the $i$-th and $j$-th coils respectively.

\begin{equation}\label{eq:SHS_blocks}
\mathbf{S}^{H}\mathbf{S} = 
\begin{pmatrix}
\mathbf{S}_1^{H}\mathbf{S}_1 & \cdots & \mathbf{S}_1^{H}\mathbf{S}_{N_c} \\
\vdots  &  \ddots & \vdots  \\
\mathbf{S}_{N_c}^{H}\mathbf{S}_1 & \cdots & \mathbf{S}_{N_c}^{H}\mathbf{S}_{N_c} 
\end{pmatrix}
\end{equation}

Since $\mathbf{S}^{H}\mathbf{S}$ is Hermitian, we only need to find the values within the lower triangle of the matrix, which are the lower triangle blocks $\mathbf{S}_i^{H}\mathbf{S}_j$ with $i>j$. For a given lower triangle block $\mathbf{S}_i^{H}\mathbf{S}_j$  ($i>j$), all of its elements are interpolated from the densely sampled Fourier transform of the dot product of the spatial domain $B_1^+$ maps of the $i$-th and $j$-th coils. The $(m,n)$-th element is the value interpolated at $(\vec{k}_m-\vec{k}_n)$ away from DC. Therefore, to solve all the columns of $\mathbf{W}$, the total number of FFT operations needed in this efficient method is only $N_c(N_c+1)/2$, instead of ${N_t}{N_c}$, and the storage size is also much smaller. In addition to the efficiency related to FFT, this algorithm saves a large matrix multiplication between $\mathbf{S}^{H}$ and $\mathbf{S}$ for the solving of every single column of $\mathbf{W}$. 
Although an alternative strategy of determining each column of $\mathbf{S}$ explicitly by interpolating the $N_s$ sample points from unshifted $B_1^+$ Fourier transforms can also reduce the number of FFT operations, it still has the problem of large multiplications between $\mathbf{S}^{H}$ and $\mathbf{S}$ matrices. Furthermore, to calculate each element of $\mathbf{S}^{H}\mathbf{S}$, there are $2N_s$ number of interpolations involved with this sub-optimal strategy. While with the efficient algorithm there is only one interpolation from the densely sampled Fourier Transform of the dot product of two spatial domain $B_1^+$ maps needed. The extra interpolations involved in the sub-optimal strategy would introduce a larger error to $\mathbf{S}^{H}\mathbf{S}$.

\subsection*{ Off-Resonance correction}
%\par As shown in Equation \ref{eq:SpatialDomainEquation} when we introduced spatial domain design in Chapter \ref{chap:SpatialDesign}, the equation connecting a pTx pulse to its excitation pattern has an off-resonance dependent term 
$e^{\imath \gamma \Delta B_0(\mathbf{x})(t-T)}$. 
In order to model the off-resonance into our k-space domain design, we adapted a fast implementation of the off-resonance compensation with time segmentation approximation and weighted least-squares interpolators \cite{fessler2005toeplitz}. 
By segmenting the RF pulse into multiple segments, this method disentangles the space-and-time dependent phase accrual induced by main field inhomogeneity into either space- or time- dependent terms, and therefore a same system matrix can be applied within each segment. Mathematically, it is
\begin{equation*}
e^{\imath \Delta\omega_j (t_i-T) }\approx\sum_{l=1}^{L} g_{l}(t_i) e^{\imath \theta_{l}(\Delta\omega_j)}
\end{equation*}
where $i=1\dots Nt$ indexes the RF samples, $j=1\dots Ns$ indexes the spatial locations, and $l$ indexes the time segments. The left term is the problematic space-and-time dependent term, with $\Delta\omega_j$ representing the off-resonance at the location-$j$ obtained from a field map and $T$ being the pulse duration. $g_{il}=b_{l}(t_i)$ are the time-dependent-only weights called temporal interpolators applied upon RF samples. $h_{lj}=e^{\imath \theta_{l}(\Delta\omega_j)}$ are the space-dependent-only phase accrual called spatial interpolators, applied upon the excitation pattern of each weighted RF segment. Therefore, the spatial domain formulation becomes
\begin{equation*}
\mathbf{d}=\sum_{l=1}^{L}diag\{\mathbf{h}_l\}\mathbf{A}(diag\{\mathbf{g}_l\}\mathbf{b})
\end{equation*}
The spatial interpolators can be absorbed into the sensitivity maps. The $\mathbf{S}$ matrix in Equation \ref{eq:MatrixForm} will then become 
\begin{equation*}
\mathbf{S_{offres}}=\sum_{l=1}^{L}\mathbf{S}_l diag\{\mathbf{g}_l\}
\end{equation*}
where the columns of the $\mathbf{S}_l$ matrix, instead of being the aforementioned k-space sensitivity maps, are now the Fourier transform of the dot products of the spatial $B_1^+$ maps and the spatial interpolatros of the $l$-th segment, shifted to corresponding excitation k-space locations. 
The efficient construction of $\mathbf{S_{offres}}^{H}\mathbf{S_{offres}}$ follows the same concept mentioned before. 
But with off-resonance, for the elements within an element matrix $\mathbf{S}_i^{H}\mathbf{S}_j$ in Equation \ref{eq:SHS_blocks}, instead of being interpolated from one densely sampled Fourier transform map, which is of the dot product of the spatial domain $B_1^+$ maps of the $i$-th and $j$-th coils, now they are interpolated from $L(L+1)/2$ different Fourier transform maps, each of which has underlying spatial $B_1^+$ maps with two sets of spatial interpolators absorbed into them. For the $(m,n)$-th element within the element matrix $\mathbf{S}_i^{H}\mathbf{S}_j$, the $L(L+1)/2$ interpolated values were then summed with weights $g_{l_i}(t_m)g_{l_j}(t_n)$, where $l_i$ and $l_j$ are the two segment numbers of the spatial interpolators observed into the corresponding $B_1^+$ maps. 
Compared to the efficient construction method without off-resonance, taking off-resonance into account increases the total number of FFT operations form $N_c(N_c+1)/2$ to $N_cL(N_cL+1)/2$, and increased the number of interpolations needed for each element from 1 to $L(L+1)/2$.

>>>>

\subsection*{k-Space-Domain Subproblems}
We propose to compute a sparse 
design matrix $\bm{W}$ that directly maps the Fourier transform of a desired spatial 
excitation pattern to a vector of RF pulses that together produce the pattern, as:
\begin{equation}
\bm{b} = \bm{W} \mathcal{F}(\bm{d}),
\end{equation}
where $\bm{b}$ is a vector of concatenated multi-channel RF pulses with length $N_t \cdot N_c$, 
where $N_t$ and $N_c$ are the numbers of time points and coils, respectively,
and $\mathcal{F}(\bm{d})$ is a length-$N_s$ vector containing the 
Fourier transform of the desired k-space excitation pattern $\bm{d}$.
$\bm{W}$ has dimensions $N_t \cdot N_c \times N_s$.
Once the entries of $\bm{W}$ are computed, 
the RF design problem can be instantaneously solved by a sparse matrix multiplication, 
without a full matrix inversion or an iterative algorithm. 
Here we describe how to break down the problem of constructing $\bm{W}$ 
into a set of smaller independent subproblems that can be solved in parallel.

%Furthermore, in design problems involved iterative procedures, such as the phase updates in magnitude-least-squares pulse designs \cite{setsompop2008magnitude} and the regularization factor updates in power/roughness constrained pulse design, the sparsity of the $\bm{W}$ matrix makes the iterative process much more efficient in terms of memory and speed. In the work of Katscher et al \cite{katscher2003transmit}, parallel transmit pulse design was also formulated in the k-space domain. However, in Ref \cite{katscher2003transmit}, the matrix $\bm{W}$ was solved as a whole, which required constructing and inverting a large $\bm{S}^H\bm{S}$ matrix, which will be explained in the following text. Our proposed method finely divides the process of finding $\bm{W}$ into many independent instances, which allows the use of parallel computing to largely accelerate the computation. Furthermore, the $\bm{S}^H\bm{S}$ matrices of different instances are each small, and our proposed method efficiently constructs these matrices, which further accelerates the computation.

\par The columns of $\bm{W}$ can be solved independently, 
based on the observation that if the desired k-space pattern is a delta function at a k-space location 
$\vec{k}_j$ ($1\leq j \leq N_s$), 
then only the $j$-th column $\bm{w}_j$ of $\bm{W}$ is required to relate 
the desired k-space pattern to the desired RF pulse. 
In this case, the vector $\bm{w}_j$ is identical to the desired RF. 
Therefore, the problem of finding the $j$-th column $\bm{w}_j$ of $\bm{W}$ can be restated as the following: 
what should the RF pulse samples be to generate a unit delta function at the location $\vec{k}_j$ in k-space, 
and zeros elsewhere? 
Mathematically, the relationship that should be satisfied by the weights in column $\bm{w}_j$  can be expressed as:   
\begin{equation}\label{eq:ScalerForm}
\delta(\vec{k}-\vec{k}_j)=\sum_{i=1}^{N_t}\sum_{c=1}^{N_c} w_c(\vec{k}_i) s_c(\vec{k}-\vec{k}_i)
\end{equation}
where $i$ indexes time points in the pulses' excitation k-space trajectory, 
and $c$ indexes coils. 
$w_c(\vec{k}_i)$ is the $((c-1)N_t+i)$-th element of the $\bm{w}_j$ vector,
and $s_c(\vec{k}-\vec{k}_i)$ is the Fourier transform of coil-$c$'s $B_1^+$ map, 
shifted to be centered at excitation k-space location $\vec{k}_i$. 
Equation \ref{eq:ScalerForm} can be restated in matrix-vector form as:                                       
\begin{equation}\label{eq:MatrixForm}
\bm{Sw}_j=\bm{\delta}_j,
\end{equation}
where $\bm{S}$ is a $N_s \times N_c \cdot N_t$ 
matrix containing values of $s_c(\vec{k}-\vec{k}_i)$ for all $c$ and $i$,
and $\bm{\delta}_j$ is a vector containing a one in the entry corresponding to excitation k-space location
$\vec{k}_j$, and zeros in all other entries. 
The regularized pseudoinverse solution for $\bm{w}_j$ is:                      
\begin{equation}\label{eq:Solution}
\bm{w}_j=\left( \bm{S}^{H} \bm{S} + \lambda \bm{I} \right) ^{-1} \bm{s}_j^{H}
\end{equation}
where $\bm{s}_j^{H}$ is the conjugate transpose of the $j$-th row of $\bm{S}$.

\par If all points on the excitation k-space trajectory are considered in Equation \ref{eq:ScalerForm},
then there is no reduction in computation compared to a conventional pseudoinverse-based pulse design.
However, because $B_1^+$ maps are compact in excitation k-space and localized near DC,
only a small number ($\ll N_t$) of trajectory points near the target location $\vec{k}_j$ can contribute significantly
to the excitation pattern at $\vec{k}_j$ or its neighbors, 
and therefore need to be considered in the subproblem;
all other trajectory points can be ignored.
This means we can modify Equation \ref{eq:ScalerForm} to:
\begin{equation}
\delta(\vec{k}-\vec{k}_j) = \sum_{\vec{k}_i \in \mathbb{N}}\sum_{c=1}^{N_c} w_c(\vec{k}_i) s_c(\vec{k}-\vec{k}_i),
\end{equation} 
where $\mathbb{N}$ is the set of trajectory points in the vicinity of $\vec{k}_j$. 
This reduced set of trajectory points reduces the column dimension of $\bm{S}$.
Since target locations that are distant to $\vec{k}_j$ can also be ignored,
the row dimension of $\bm{S}$ is also reduced to the size of a smaller neighborhood immediately surrounding $\vec{k}_j$. 
These reduced problem dimensions further imply that the design matrix $\bm{W}$ is sparse. 
Figure \ref{fig:Patch}a illustrates a target point $\vec{k}_j$, 
the nearby points on a spiral trajectory that need to be included in the solution for $\bm{w}_j$,
and the neighboring target location points where a zero excitation must be enforced. 
In the following, we define target and trajectory neighborhood sizes, and describe an efficient 
method to construct the $\bm{S}^{H}\bm{S}$ matrices.

@@

\par A key advantage of the proposed pulse design formalism is that it can be broken down into many small problems and finely parallelized, because the process of solving each column of $\bm{W}$ is independent of other columns. The columns can be solved fully independently, or patchwise for target locations in neighborhoods, wherein all points in the neighborhood share the same $\bm{S}^{H}\bm{S}$ matrix. 

@@

An example of this concept is shown in Figure \ref{fig:Patch}. For simplicity of visualization, it is demonstrated in a 2D problem, but the concept is easily extended to 3D. The figure depicts the solution of the $\bm{W}$ matrix 16 columns at a time, whose corresponding delta functions live within a patch of width 4 cycle/FOV (yellow points in Figure \ref{fig:Patch}). In this instance where the inclusion width is also 4 cycle/FOV, we truncate the $B_1^+$ map Fourier transforms to zero outside a 4 by 4 cycle/FOV square centered at DC. As a result, we must consider all the excitation trajectory points within a 12 by 12 cycle/FOV square, which is 4 cycle/FOV extended outside of the patch width 4 square, since they may contribute energy to the target locations corresponding to the 16 $\bm{W}$ matrix columns. For a single instance of the 16-point target patch shown in Figure \ref{fig:Patch}, the 46 green points on the excitation trajectory may contribute significant energy and are considered in the weight design. Furthermore, we must constrain the total excitation to zero at all k-space locations within a 20 by 20 cycle/FOV red patch in Figure \ref{fig:Patch}. This is because any energy deposited at each considered excitation trajectory points will affect not only the target points in the yellow square, but also the whole 2$\times$4 wide patch around that trajectory point. And the 20 by 20 cycle/FOV red patch is the total of the points that can be affected by all the excitation trajectory points considered in this instance. Therefore, in the solution for this instance, $\bm{w}$ is a 46$N_c\times$16 matrix, $\bm{S}$ is a 400$N_c\times$46$N_c$ matrix, $\bm{S}^{H}\bm{S}$ is a 46$N_c\times$46$N_c$ matrix, and $\bm{s}_i^{H}$ is a 46$N_c\times$16 matrix. 
%Since $\bm{s}_i^{H}$ is no longer related to the $i$th row of $\bm{S}$ and $i$th column of $\bm{W}$ only, it will be denoted as $\bm{S_{targ}}$ in later discussions. 

@@

\par The patch width determines the total number of instances to be solved. The inclusion width determines the accuracy of $B_1^+$ information utilized in the design. Together, the patch width and the inclusion width determine how many excitation trajectory points are included in each instance, which further determines the size of $\bm{S}^{H}\bm{S}$ whose construction is most computationally burdensome within an instance. Therefore, increasing patch width reduces the number of $\bm{S}^{H}\bm{S}$ matrices to be constructed but increases the burden of constructing each of them. An optimal patch width should be decided case by case for different excitation trajectories.

\subsection*{Efficient construction of $\bm{S}^{H}\bm{S}$ matrix}
\par Constructing $\bm{S}^{H}\bm{S}$ by finding $\bm{S}$ is computationally expensive. As described previously, the $((c-1)N_c+t)$-th column of the  $\bm{S}$ matrix, $\bm{S}_c(\vec{k}-\vec{k}_t)$, is the k-space sensitivity map of the $c$-th coil $\bm{S}_c(\vec{k})$ shifted to excitation k-space trajectory location $\vec{k}_t$. These shifts for each column of $\bm{S}$ are achieved by applying a linear phase modulation to the spatial domain $B_1^+$ map, followed by a fast Fourier transform (FFT). This process becomes computationally overwhelming when there are many RF samples, which is often the case. Furthermore, the matrix multiplication between $\bm{S}^{H}$ and $\bm{S}$ is also computationally intensive due to the large row dimension of $\bm{S}$ (400$Nt$ in the above example in Figure \ref{fig:Patch}).

\par Inspired by a rapid GRAPPA calibration method for arbitrary 2D/3D non-Cartesian trajectories proposed in Ref \cite{luo2019grappa}, we have developed a fast and memory-efficient algorithm to solve this problem. A key innovation of Ref \cite{luo2019grappa} is that the algorithm constructs the Hermitian $\bm{S}^{H}\bm{S}$ matrix directly by finding each of the matrix's lower triangle elements through interpolation, instead of building the $\bm{S}$ matrix explicitly followed by matrix multiplication. Conceptually, each element in the Hermitian matrix $\bm{S}^{H}\bm{S}$ is the vector sum of two shifted k-space sensitivity maps, which is also one single point of their convolution. With the multiplication property of Fourier transform, this convolution in k-space can be calculated from the Fourier transform of the dot product of the two spatial domain $B_1^+$ maps. For a more detailed derivation, please refer to Ref \cite{luo2019grappa}.
In practice, $\bm{S}^{H}\bm{S}$ is constructed in the following way. We first represent the $\bm{S}^{H}\bm{S}$ matrix with block matrices. Each block matrix $\bm{S}_i^{H}\bm{S}_j$ is of size $N_n$ by $N_n$, and $\bm{S}_i$ and $\bm{S}_j$ are the columns of $\bm{S}$ related to the k-space sensitivity maps of the $i$-th and $j$-th coils respectively.

\begin{equation}\label{eq:SHS_blocks}
\bm{S}^{H}\bm{S} = 
\begin{pmatrix}
\bm{S}_1^{H}\bm{S}_1 & \cdots & \bm{S}_1^{H}\bm{S}_{N_c} \\
\vdots  &  \ddots & \vdots  \\
\bm{S}_{N_c}^{H}\bm{S}_1 & \cdots & \bm{S}_{N_c}^{H}\bm{S}_{N_c} 
\end{pmatrix}
\end{equation}

Since $\bm{S}^{H}\bm{S}$ is Hermitian, we only need to find the values within the lower triangle of the matrix, which are the lower triangle blocks $\bm{S}_i^{H}\bm{S}_j$ with $i>j$. For a given lower triangle block $\bm{S}_i^{H}\bm{S}_j$  ($i>j$), all of its elements are interpolated from the densely sampled Fourier transform of the dot product of the spatial domain $B_1^+$ maps of the $i$-th and $j$-th coils. The $(m,n)$-th element is the value interpolated at $(\vec{k}_m-\vec{k}_n)$ away from DC. Therefore, to solve all the columns of $\bm{W}$, the total number of FFT operations needed in this efficient method is only $N_c(N_c+1)/2$, instead of ${N_t}{N_c}$, and the storage size is also much smaller. In addition to the efficiency related to FFT, this algorithm saves a large matrix multiplication between $\bm{S}^{H}$ and $\bm{S}$ for the solving of every single column of $\bm{W}$. 
Although an alternative strategy of determining each column of $\bm{S}$ explicitly by interpolating the $N_s$ sample points from unshifted $B_1^+$ Fourier transforms can also reduce the number of FFT operations, it still has the problem of large multiplications between $\bm{S}^{H}$ and $\bm{S}$ matrices. Furthermore, to calculate each element of $\bm{S}^{H}\bm{S}$, there are $2N_s$ number of interpolations involved with this sub-optimal strategy. While with the efficient algorithm there is only one interpolation from the densely sampled Fourier Transform of the dot product of two spatial domain $B_1^+$ maps needed. The extra interpolations involved in the sub-optimal strategy would introduce a larger error to $\bm{S}^{H}\bm{S}$.

\subsection*{ Off-Resonance correction}
%\par As shown in Equation \ref{eq:SpatialDomainEquation} when we introduced spatial domain design in Chapter \ref{chap:SpatialDesign}, the equation connecting a pTx pulse to its excitation pattern has an off-resonance dependent term 
$e^{\imath \gamma \Delta B_0(\bm{x})(t-T)}$. 
In order to model the off-resonance into our k-space domain design, we adapted a fast implementation of the off-resonance compensation with time segmentation approximation and weighted least-squares interpolators \cite{fessler2005toeplitz}. 
By segmenting the RF pulse into multiple segments, this method disentangles the space-and-time dependent phase accrual induced by main field inhomogeneity into either space- or time- dependent terms, and therefore a same system matrix can be applied within each segment. Mathematically, it is
\begin{equation*}
e^{\imath \Delta\omega_j (t_i-T) }\approx\sum_{l=1}^{L} g_{l}(t_i) e^{\imath \theta_{l}(\Delta\omega_j)}
\end{equation*}
where $i=1\dots Nt$ indexes the RF samples, $j=1\dots Ns$ indexes the spatial locations, and $l$ indexes the time segments. The left term is the problematic space-and-time dependent term, with $\Delta\omega_j$ representing the off-resonance at the location-$j$ obtained from a field map and $T$ being the pulse duration. $g_{il}=b_{l}(t_i)$ are the time-dependent-only weights called temporal interpolators applied upon RF samples. $h_{lj}=e^{\imath \theta_{l}(\Delta\omega_j)}$ are the space-dependent-only phase accrual called spatial interpolators, applied upon the excitation pattern of each weighted RF segment. Therefore, the spatial domain formulation becomes
\begin{equation*}
\bm{d}=\sum_{l=1}^{L}diag\{\bm{h}_l\}\bm{A}(diag\{\bm{g}_l\}\bm{b})
\end{equation*}
The spatial interpolators can be absorbed into the sensitivity maps. The $\bm{S}$ matrix in Equation \ref{eq:MatrixForm} will then become 
\begin{equation*}
\bm{S_{offres}}=\sum_{l=1}^{L}\bm{S}_l diag\{\bm{g}_l\}
\end{equation*}
where the columns of the $\bm{S}_l$ matrix, instead of being the aforementioned k-space sensitivity maps, are now the Fourier transform of the dot products of the spatial $B_1^+$ maps and the spatial interpolatros of the $l$-th segment, shifted to corresponding excitation k-space locations. 
The efficient construction of $\bm{S_{offres}}^{H}\bm{S_{offres}}$ follows the same concept mentioned before. 
But with off-resonance, for the elements within an element matrix $\bm{S}_i^{H}\bm{S}_j$ in Equation \ref{eq:SHS_blocks}, instead of being interpolated from one densely sampled Fourier transform map, which is of the dot product of the spatial domain $B_1^+$ maps of the $i$-th and $j$-th coils, now they are interpolated from $L(L+1)/2$ different Fourier transform maps, each of which has underlying spatial $B_1^+$ maps with two sets of spatial interpolators absorbed into them. For the $(m,n)$-th element within the element matrix $\bm{S}_i^{H}\bm{S}_j$, the $L(L+1)/2$ interpolated values were then summed with weights $g_{l_i}(t_m)g_{l_j}(t_n)$, where $l_i$ and $l_j$ are the two segment numbers of the spatial interpolators observed into the corresponding $B_1^+$ maps. 
Compared to the efficient construction method without off-resonance, taking off-resonance into account increases the total number of FFT operations form $N_c(N_c+1)/2$ to $N_cL(N_cL+1)/2$, and increased the number of interpolations needed for each element from 1 to $L(L+1)/2$.

### 2020-08-17 Manuscript/kPtx_paper.tex
wag edited first section of theory

\title{k-Space Domain Parallel Transmit Pulse Design}
\author{Jun Ma$^{1,2}$ and William A. Grissom$^{*1,2}$}
\maketitle
\begin{flushleft}
$^1$Vanderbilt University Institute of Imaging Science, Nashville, TN, United States\\
$^2$Department of Biomedical Engineering, Vanderbilt University, Nashville, TN, United States\\

>>>>

\title{k-Space Domain Parallel Transmit Pulse Design}
\author{Jun Ma$^{1,2}$, Bernhard Gruber, Xinqiang Yan$^{1,3}$, and William A. Grissom$^{1,2*}$}
\maketitle
\begin{flushleft}
$^1$Vanderbilt University Institute of Imaging Science, Nashville, TN, United States\\
$^2$Department of Biomedical Engineering, Vanderbilt University, Nashville, TN, United States\\
$^3$Department of Radiology and Radiological Sciences, Vanderbilt University, Nashville, TN, United States\\
